What Doubling Tricks Can and Cannot Do for Multi-Armed Bandits

### 2019-03-29 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Okay, okay, apparently « on the one hand » is (slightly) more frequent. https://dictionary.cambridge.org/fr/dictionnaire/anglais/on-the-one-hand-on-the-other-hand

Like before in this chapter, we consider an LPWA network composed of a gateway and a large number of end-devices that regularly send short data packets, where $K$ channels ($K>1$) are available for the transmission of their packets.
%
We assume that this network is constituted by two types of devices:
on one hand, we have \emph{static} devices that operate in one channel\footnote{~Note that, for unlicensed bands, this definition also encompasses any device following a different standard or trying to establish communication with gateways of other networks.} in order to communicate with the gateway.
%
On the other hand, there are  IoT devices, that possess the additional advantage of being able to select any of the $K$ available channels to perform their transmissions.

@@

We interpret this fact in two different situations.
\begin{itemize}
	\item
	On one hand, in a congested network, when devices suffer from a large probability of collision on their first transmission (\ie, $p_c$ is not so small), then $p_{c1}\simeq p_c$ and so devices cannot really hope to reduce their collision probabilities even if the use a different channel for retransmission.
	\item
	On the other hand, if $p_c$ is small enough, \ie, in a network not yet too congested, then our derivation shows that $p_{c1} > p_c$, meaning that the possible gain of retransmitting in a different channel that the one used for the first transmission can be large, in terms of collision probability (\eg, up-to $10\%$ in this experimental setting).
	In other words, when learning can be useful (small $p_c$), learning to retransmit in a different channel can have a large impact on the global collision rate,
	thus justifying our approach.
\end{itemize}

>>>>

Like before in this chapter, we consider an LPWA network composed of a gateway and a large number of end-devices that regularly send short data packets, where $K$ channels ($K>1$) are available for the transmission of their packets.
%
We assume that this network is constituted by two types of devices:
On the one hand, we have \emph{static} devices that operate in one channel\footnote{~Note that, for unlicensed bands, this definition also encompasses any device following a different standard or trying to establish communication with gateways of other networks.} in order to communicate with the gateway.
%
On the other hand, there are  IoT devices, that possess the additional advantage of being able to select any of the $K$ available channels to perform their transmissions.

@@

We interpret this fact in two different situations.
\begin{itemize}
	\item
	On the one hand, in a congested network, when devices suffer from a large probability of collision on their first transmission (\ie, $p_c$ is not so small), then $p_{c1}\simeq p_c$ and so devices cannot really hope to reduce their collision probabilities even if the use a different channel for retransmission.
	\item
	On the other hand, if $p_c$ is small enough, \ie, in a network not yet too congested, then our derivation shows that $p_{c1} > p_c$, meaning that the possible gain of retransmitting in a different channel that the one used for the first transmission can be large, in terms of collision probability (\eg, up-to $10\%$ in this experimental setting).
	In other words, when learning can be useful (small $p_c$), learning to retransmit in a different channel can have a large impact on the global collision rate,
	thus justifying our approach.
\end{itemize}

### 2019-03-29 2-Chapters/4-Chapter/chapter4.tex
Okay, okay, apparently « on the one hand » is (slightly) more frequent. https://dictionary.cambridge.org/fr/dictionnaire/anglais/on-the-one-hand-on-the-other-hand

To sum-up, we focussed in this chapter on models of IoT networks, and we proposed to use classical stationary multi-armed bandit learning algorithms implemented in a selfish and decentralized manner by each of the dynamic devices of the IoT network.
We presented two models of wireless IoT network without sensing, inspired by the ALOHA protocol, and with or without retransmission of uplink packages in case of collisions.
In both cases, we try to model the existing standards, like the LoRa standard, and we demonstrated the efficiency of the proposed MAB-based approach in both numerical simulations and empirical measurements on real wireless radio signals.
%
On one hand, a satisfying message is that this learning-based approach seems to be very efficient, as it allows the IoT devices to automatically and independently increase their successful transmission rate.
On the other hand, it is quite surprising that stochastic MAB algorithms can be of any use in such non-stationary applications.
Sadly, analyzing our model with thousands of independent devices communicating and learning in their own (random) time scale turned out to be of extreme difficulty.

>>>>

To sum-up, we focussed in this chapter on models of IoT networks, and we proposed to use classical stationary multi-armed bandit learning algorithms implemented in a selfish and decentralized manner by each of the dynamic devices of the IoT network.
We presented two models of wireless IoT network without sensing, inspired by the ALOHA protocol, and with or without retransmission of uplink packages in case of collisions.
In both cases, we try to model the existing standards, like the LoRa standard, and we demonstrated the efficiency of the proposed MAB-based approach in both numerical simulations and empirical measurements on real wireless radio signals.
%
On the one hand, a satisfying message is that this learning-based approach seems to be very efficient, as it allows the IoT devices to automatically and independently increase their successful transmission rate.
On the other hand, it is quite surprising that stochastic MAB algorithms can be of any use in such non-stationary applications.
Sadly, analyzing our model with thousands of independent devices communicating and learning in their own (random) time scale turned out to be of extreme difficulty.

### 2019-03-29 3-Appendices/2-Appendix/Revisiting_Doubling_Trick_for_Multi-Armed_Bandits.git/Revisiting_Doubling_Trick_for_Multi-Armed_Bandits.tex
Okay, okay, apparently « on the one hand » is (slightly) more frequent. https://dictionary.cambridge.org/fr/dictionnaire/anglais/on-the-one-hand-on-the-other-hand

The \MOSS{} algorithm, proposed by \cite{Audibert2009minimax}, is the first stochastic bandit algorithm
to enjoy a problem-dependent logarithmic regret
and to be optimal in a \emph{minimax} sense, as its regret is proved to be upper bounded by $\sqrt{KT}$, for bandit models with rewards in $[0,1]$.
However the corresponding constant $C(\bm \nu)$ is proportional to $K/\Delta_{\min}$, where $\Delta_{\min} = \min_{k} (\mu^*-\mu_k)$ is the minimal gap, which worsen the constant of \UCBone{}.
Another drawback of \MOSS{} is that it is \emph{not} anytime.
%
These two shortcoming have been overcame recently in two different works.
On one hand, the \MOSSAnytime{} algorithm \citep{Degenne16} is minimax optimal and anytime, but its problem-dependent regret does not improve that of \MOSS.
On the other hand, the \KLUCBpp{} algorithm \citep{Menard17} is simultaneously minimax optimal and asymptotically optimal (\ie, it has the best problem-dependent constant $C(\bm\nu)$), but it is not anytime. A natural question is thus to know whether a Doubling Trick could overcome this limitation.

@@

For a fixed $\gamma$ and $\delta$, minimizing $\ell(\gamma, \delta, T_0, b)$ does not always give a unique solution.
On one hand, if $\gamma \gtrsim 0.384$, there is a unique solution $b^*(\gamma) > 1$ minimizing the $\frac{b^{\gamma} (b-1)^{\gamma}}{b^{\gamma} - 1}$ term, solution of $b^{\gamma+1} - 2b + 1 = 0$, but without a closed form if $\gamma$ is unknown.
%
On the other hand, for any $\gamma$, the term depending on $\delta$ tends quickly to $1$ when $T_0$ increases.

>>>>

The \MOSS{} algorithm, proposed by \cite{Audibert2009minimax}, is the first stochastic bandit algorithm
to enjoy a problem-dependent logarithmic regret
and to be optimal in a \emph{minimax} sense, as its regret is proved to be upper bounded by $\sqrt{KT}$, for bandit models with rewards in $[0,1]$.
However the corresponding constant $C(\bm \nu)$ is proportional to $K/\Delta_{\min}$, where $\Delta_{\min} = \min_{k} (\mu^*-\mu_k)$ is the minimal gap, which worsen the constant of \UCBone{}.
Another drawback of \MOSS{} is that it is \emph{not} anytime.
%
These two shortcoming have been overcame recently in two different works.
On the one hand, the \MOSSAnytime{} algorithm \citep{Degenne16} is minimax optimal and anytime, but its problem-dependent regret does not improve that of \MOSS.
On the other hand, the \KLUCBpp{} algorithm \citep{Menard17} is simultaneously minimax optimal and asymptotically optimal (\ie, it has the best problem-dependent constant $C(\bm\nu)$), but it is not anytime. A natural question is thus to know whether a Doubling Trick could overcome this limitation.

@@

For a fixed $\gamma$ and $\delta$, minimizing $\ell(\gamma, \delta, T_0, b)$ does not always give a unique solution.
On the one hand, if $\gamma \gtrsim 0.384$, there is a unique solution $b^*(\gamma) > 1$ minimizing the $\frac{b^{\gamma} (b-1)^{\gamma}}{b^{\gamma} - 1}$ term, solution of $b^{\gamma+1} - 2b + 1 = 0$, but without a closed form if $\gamma$ is unknown.
%
On the other hand, for any $\gamma$, the term depending on $\delta$ tends quickly to $1$ when $T_0$ increases.
